---
title: "accidents"
output: html_document
date: '2022-05-08'
---

Install Packages

```{r}
install.packages("plotly")
install.packages('maps')
install.packages('usmap')
```


Importing Libraries

```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
library(plotly)
library(usmap)
library(choroplethr)
library(choroplethrMaps)
library(corrplot)
```

View Data

```{r}
View(FilteredData)

dim(FilteredData)

```

DATA CLEANING

Cleaning Date Column and converting it into Year, Month and Day


```{r}

FilteredData$Crash.Date <- as.Date(FilteredData$Crash.Date, "%m/%d/%Y")

FilteredData <- tidyr::separate(FilteredData, 
                                Crash.Date, c('Year', 'Month', 'Day'),
                                sep = "-",remove = FALSE)

```


Removing columns with more missing values

```{r}

accidents <- FilteredData[,!(colnames(FilteredData) %in% 
                               c('Distance.To.Cross.Street', 
                                  'MilePost',                                                         'Posted.Speed.Cross.Street',                                        'Driver.Physical.Status',                                           'month', 'year','day'))]

accidents <- accidents[, !(colnames(accidents) %in%
                            c('Alcohol.Test.Results',
                                  'Driver.Physical.Status',
                                  'Driver.Zip.Code',
                                  'Temporary.Traffic.Control.Zone',
                                  'Route '))]

```


Dropping rows with missing value

```{r}
accidents <- accidents %>% drop_na(Latitude, Longitude)
```

Converting Time into Hours and Minutes

```{r}

accidents <- accidents%>%
  separate(Crash.Time, into = c("hours", "minutes"))

accidents$hours <- as.numeric(accidents$hours)

accidents$hours[is.na(accidents$hours)]<-median(accidents$hours,na.rm=TRUE)


```

Converting columns from character to numeric

```{r}
accidents$Route[is.na(accidents$Route)]<-mean(accidents$Route,na.rm=TRUE)
accidents$Road.Character[is.na(accidents$Road.Character)]<-mean(accidents$Road.Character,na.rm=TRUE)
accidents$Surface.Condition[is.na(accidents$Surface.Condition)]<-mean(accidents$Surface.Condition,na.rm=TRUE)
accidents$ Environmental.Condition[is.na(accidents$ Environmental.Condition)]<-mean(accidents$ Environmental.Condition,na.rm=TRUE)
accidents$Temporary.Traffic.Control.Zone[is.na(accidents$Temporary.Traffic.Control.Zone)]<-mean(accidents$Temporary.Traffic.Control.Zone,na.rm=TRUE)
accidents$Road.Surface.Type[is.na(accidents$Road.Surface.Type)]<-mean(accidents$Road.Surface.Type,na.rm=TRUE)
accidents$Light.Condition[is.na(accidents$Light.Condition)]<-mean(accidents$Light.Condition,na.rm=TRUE)
accidents$Road.Divided.By[is.na(accidents$Road.Divided.By)]<-mean(accidents$Road.Divided.By,na.rm=TRUE)

```

Adding column state


```{r}
accidents$state <- 'NJ'
```





Number of accidents in a year

```{r}
total_acci <- accidents %>%
              group_by(Year)%>%
              summarise(count_by_year = n())
total_acci
```


```{r}
accidents%>%
  group_by(Month)%>%
  summarise(accidents = cumsum(n()))%>%
  ggplot(aes(Month, accidents))+
  geom_smooth()
```


Visualizing number of accidents per month

```{r}
accidents%>%
  group_by(Month)%>%
  summarise(accidents = n())%>%
  ggplot(aes(Month, accidents))+
  geom_line(group = 1, color = 'red')+
  geom_point(color = 'red')+
  ggtitle('Number of Accidents per Month')
```

Visualizing number of accidents by day

```{r}
accidents%>%
  group_by(Day)%>%
  summarise(count_by_day = n())%>%
  ggplot(aes(Day, count_by_day))+
  geom_line(group = 1, color = 'red')+
  geom_point(color = 'red')+
  ggtitle('Number of Accidents by Day')
```

Accidents on Week of the Day

```{r}
accidents%>%
  group_by(Crash.Day.Of.Week)%>%
  summarise(accidents_by_week = n())%>%
  ggplot(aes(Crash.Day.Of.Week, accidents_by_week))+
  geom_line(group = 1, color = 'red')+
  geom_point(color = 'red')+
  ggtitle('Accidents vs Weekday')

```


```{r}
accidents%>%
  group_by(hours)%>%
  summarise(total_accidents = n())%>%
  ggplot(aes(x = hours, y = total_accidents))+
  geom_bar(stat = 'identity')
```

Accidents vs Alcohol Involved

```{r}
alcohol_count <- accidents %>%
                   group_by(Month, Alcohol.Involved)%>%
                   summarise(alcohol_count = n())

ggplot(data = alcohol_count, aes(Month, alcohol_count)) +
              geom_line(group = 1, color = 'blue') +
              geom_point(color = 'blue')+
              facet_wrap(~ Alcohol.Involved, scales = 'free_y', nrow = 2)+
             ggtitle('Monthly Accidents based on Alcohol Consumption')
```

Accidents with alcohol involved vs Week day



```{r}
accidents%>%
  group_by(Crash.Day.Of.Week, Alcohol.Involved)%>%
  summarise(alcohol_involved = n())%>%
  ggplot(aes(Crash.Day.Of.Week, alcohol_involved))+
  geom_line(group = 1, color = 'red')+
  geom_point(color = 'red')+
  facet_wrap(~Alcohol.Involved, scales = 'free_y', nrow = 2)+
  ggtitle('Weekly Accidents vs Alcohol Involved')

```



```{r}

accidents%>%
  group_by(Month,Crash.Day.Of.Week)%>%
  summarise(accident_count = n())%>%
  ggplot(aes(Crash.Day.Of.Week, accident_count))+
  geom_line(group = 1, color = 'red')+
  geom_point(color = 'red')+
  facet_wrap(~ Month, scale = 'free_x')+
  ggtitle('Number of Accidents in Weekday for each Month')

```


Injured vs Month

```{r}
accidents <- accidents %>%
             mutate(injured = ifelse(
                     accidents$Total.Injured == 0, "No", "Yes"))


injured <- accidents%>%
            group_by(Month, injured)%>%
            summarise(injured_count = n())
  


ggplot(data = injured, aes(Month, injured_count)) +
  geom_line(group = 1, color = 'blue') +
  geom_point(color = 'blue')+
  facet_wrap(~ injured, scales = 'free_y', nrow = 2)+
  ggtitle('Accident Injured vs Month')

```

Injured for Week of the Day

```{r}
accidents%>%
  group_by(Crash.Day.Of.Week, injured)%>%
  summarise(injured_count = n())%>%
  ggplot(aes(Crash.Day.Of.Week, injured_count))+
  geom_line(group = 1, color = 'red')+
  geom_point(color = 'red')+
  facet_wrap(~ injured, scale = 'free_y')+
  ggtitle('Total Injured vs Weekday')

```


Total killed vs Month

```{r}
accidents%>%
  group_by(Month, Total.Killed)%>%
  summarise(killed = n())%>%
  ggplot(aes(Month, killed))+
  geom_line(group = 1, color = 'red')+
  geom_point(color = 'red')+
  facet_wrap(~ Total.Killed, scale = 'free_y')+
  ggtitle('Number of People Killed for each Month')
```


Accidents vs Crash code

```{r}
accidents$Crash.Type.Code <- gsub("\\..*","",accidents$Crash.Type.Code)

crash_code <- accidents%>%
  group_by(Crash.Type.Code)%>%
  summarise(crash_code_count = n())%>%
  arrange(desc(crash_code_count))%>%
  slice(1:5)

crash_code %>%
  ggplot(aes(Crash.Type.Code, crash_code_count, fill = 'crash_code_count'))+
  geom_bar(stat = 'identity')

```


Accidents based on Gender

```{r}
accidents$Driver.Sex = str_trim(accidents$Driver.Sex) 
sum(is.na(accidents$Driver.Sex))

gender <- accidents %>% 
  group_by(Month,Driver.Sex)%>%
  filter(Driver.Sex != ' ')%>%
  summarise(count =n())


ggplot(data = gender, aes(Month, count)) +
  geom_line(group = 1, color = 'blue') +
  geom_point(color = 'blue')+
 facet_wrap(~ Driver.Sex, scales = 'free_y', nrow = 2)+
  ggtitle('Number of Accidents by Gender for each Month')
```

Accidents based on Gender for week Day

```{r}
gender_week <-  accidents %>% 
               group_by(Crash.Day.Of.Week,Driver.Sex)%>%
               filter(Driver.Sex != ' ')%>%
               summarise(accident_count =n())

ggplot(data = gender_week, aes(Crash.Day.Of.Week, accident_count)) +
  geom_line(group = 1, color = 'blue') +
  geom_point(color = 'blue')+
  facet_wrap(~ Driver.Sex, scales = 'free_y', nrow = 2)+
  ggtitle('Number of Accidents by Gender for Week Days')


```

Map view of number of accidents

```{r}
accidents$Longitude <- abs(accidents$Longitude)
accidents$Longitude <- accidents$Longitude * -1

accidents$County.Name <- tolower(accidents$County.Name)

county_map_data<-map_data("county")
state_map <- map_data("state")
nj_df <- subset(state_map, region == "new jersey")

nj_base <- ggplot(data = nj_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  theme_void()+
  geom_polygon(color = "white", fill = NA)

nj_base + geom_point(data = accidents, aes(x = Longitude, y = Latitude), inherit.aes = FALSE, color = 'red', size = 0.01)


```

Correlation Map

```{r}
correlation = cor(select(accidents, hours, Total.Killed, Total.Injured, Light.Condition, Total.Vehicles.Involved))


corrplot(correlation, type = "lower", 
         tl.col = "red", tl.srt = 90, title = 'Correlation Map')
```

```{r}

```


```{r}
colSums(is.na(FilteredData))

```


```{r}
```


```{r}
```


```{r}
View(FilteredData)

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
