---
title: "us_accidents"
output: html_document
date: '2022-05-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing Libraries

```{r cars}
library(tidyverse)
library(lubridate)
library(dplyr)
library(plotly)
library(usmap)
library(choroplethr)
library(choroplethrMaps)
library(corrplot)
```

##Installing Required Packages 


```{r}
install.packages('njtr1')
install.packages("devtools")

```


```{r}
devtools::install_github("gavinrozzi/njtr1")


```


```{r}
library(njtr1)
```


##Loading Accidents dataset

```{r}
acc_17 <- get_njtr1(year = 2017, type = "Accidents")

acc_18 <- get_njtr1(year = 2018, type = "Accidents")

acc_19 <- get_njtr1(year = 2019, type = "Accidents")

acc_20 <- get_njtr1(year = 2020, type = "Accidents")
```

##Merging dataframes

```{r}
acc_17_18 <- rbind(acc_17, acc_18)


acc_18_19 <- rbind(acc_17_18, acc_19)

accidents_df <- rbind(acc_18_19, acc_20) 
```


##Loading  drivers dataset

```{r}
dri_17 <- get_njtr1(year = 2017, type = "Drivers")


dri_18 <- get_njtr1(year = 2018, type = "Drivers")


dri_19 <- get_njtr1(year = 2019, type = "Drivers")


dri_20 <- get_njtr1(year = 2020, type = "Drivers")
```

##Merging Drivers dataframe

```{r}
dri_17_18 <- rbind(dri_17, dri_18)

dri_18_19 <- rbind(dri_17_18, dri_19)

dri <- rbind(dri_18_19, dri_20)
```

##Subsetting drivers dataframe

```{r}
driver <- dri[,c('id', 'driver_dob', 'driver_sex')]
```

##Imputing missing value in drivers dataframe

```{r}
colSums(is.na(driver))

driver <- na.omit(driver)
```

##Merging accidents df with drivers subset df

```{r}
accidents_df <- merge(accidents_df, driver, by = 'id')
```


##Dropping unwanted columns

```{r}
drops <- c("police_station","location_direction", 'route_suffix', 'sri_std_rte_identifier', 'road_character', 'unit_of_measurement', 'direction from cross street', 'is_ramp', 'ramp_tofrom_route_name', 'ramp_tofrom_route_direction', 'other_property_damage', 'distance_to_cross_street')
accidents_df = accidents_df[ , !(names(accidents_df) %in% drops)]
```

##Checking for missing values

```{r}
colSums(is.na(accidents_df))
```

##Imputing missing value for latitude and logitude based on the municipality

```{r}
accidents_df$latitude<-ave(accidents_df$latitude,accidents_df$municipality_name,FUN=function(x) 
  ifelse(is.na(x), mean(x,na.rm=TRUE), x))



accidents_df$longitude<-ave(accidents_df$longitude,accidents_df$municipality_name,FUN=function(x) 
  ifelse(is.na(x), mean(x,na.rm=TRUE), x))


```

#Imputing other missing values by mean


```{r}
for(i in 1:ncol(accidents_df)){
  accidents_df[is.na(accidents_df[,i]), i] <- median(accidents_df[,i], na.rm = TRUE)
}
```

##Creating new column year, month, day

```{r}
accidents_df$year <- format(accidents_df$crash_date, format="%Y")
accidents_df$month <- format(accidents_df$crash_date, format="%m")
accidents_df$day <- format(accidents_df$crash_date, format="%d")
```

##Creating new columns hours and minutes

```{r}
accidents_df$crash_time <- substr(as.POSIXct(sprintf("%04.0f", accidents_df$crash_time), format='%H%M'), 12, 16)


accidents_df <- accidents_df%>%
  separate(crash_time, into = c("hours", "minutes"))
```

##Converting hours from charecter to numeric

```{r}
accidents_df$hours <- as.numeric(accidents_df$hours)
```

##Visualizing county level number of accidents

```{r}
county_level_accident <- accidents_df%>%
                            group_by(county_name)%>%
                            dplyr::summarise(total_accidents=n())%>%
                            mutate(county =   factor(county_name,level=county_name))


county_level_accident <- subset(county_level_accident, 
                                county_name %in% 
                                c('ATLANTIC','BERGEN', 'BURLINGTON',                                       'CAMDEN', 'CAPE MAY', 'CUMBERLAND',                                      'ESSEX','GLOUCESTER','HUDSON',                                           'HUNTERDON', 'MERCER', 
                                  'MIDDLESEX', 'MONMOUTH', 'MORRIS',                                       'OCEAN', 'PASSAIC', 'SALEM',                                             'SOMERSET', 'SUSSEX', 	
                                  'UNION', 'WARREN'))

ggplot(county_level_accident, 
       aes(total_accidents,                                                     reorder(county_name,total_accidents),
       fill=county_name))+
geom_bar(stat='identity')+
ggtitle('Number of Accidents per County')
```

**The above graph shows more number of accidents occur in Bergen, Middlesex, Essex county**


##Visualizing number of accidents for each year

```{r}
yearly_accident <- accidents_df %>%
                      group_by(year)%>%
                      dplyr::summarise(Accident_by_year=n())

ggplot(yearly_accident, 
       aes(year, 
           Accident_by_year, 
           fill=year))+
geom_bar(stat='identity')+
ggtitle('Total Number of Accidents by Year')
```
**The above graph shows there is decrease in total number of accidents in 2020 compared toprevious years**


##Visualizing number of accidents by month

```{r}
accidents_df%>%
  group_by(year, month)%>%
  dplyr::summarise(accidents=n())%>%
  ggplot(aes(month, accidents, group=year, color =year))+
  geom_line(size=1)+
  ggtitle('Total Accidents for each month and year')

```
**There is increase in number of cases during the end of yearcompare to other months**


##Visualizing number of accidents by week

```{r}
weekly_count <- accidents_df%>%
                   group_by(crash_day_of_week)%>%
                   dplyr::summarise(Number_of_Accidents=n())


weekly_count <- weekly_count[-c(7), ]


```

```{r}
weekly_count$crash_day_of_week <- ordered(weekly_count$crash_day_of_week,                                            levels=c("MO", "TU", "WE",                                                        "TH", "FR", "SA",                                                        "SU"))

ggplot(weekly_count, 
       aes(x = crash_day_of_week,
           y=Number_of_Accidents))+
geom_segment( aes(x=crash_day_of_week,
                  xend=crash_day_of_week, 
                  y=0,                         
                  yend=Number_of_Accidents))+
geom_point( size=5, 
            color="red", 
            fill=alpha("orange", 0.3), 
            alpha=0.7,
            shape=21, stroke=2)+
ggtitle('Number of Accidents by Weekday')
```

##Visualizing weekend trend for each hour

```{r}
accident_weekends <- subset(accidents_df,
                           accidents_df$crash_day_of_week == "SA" |                                 accidents_df$crash_day_of_week == "SU")%>%
                   group_by(hours)%>%
                   dplyr::summarise(total_accident=n())

ggplot(accident_weekends, 
       aes(hours, 
           total_accident, 
           color='red'))+
geom_point()+
geom_line(color='red')+
ggtitle('Number of Accidents on Weekends')

```

#Visualizing weekday trend by hours

```{r}
accident_weekdays <- subset(accidents_df,
                           accidents_df$crash_day_of_week != "SA" |                                 accidents_df$crash_day_of_week != "SU")%>%
                   group_by(hours)%>%
                   dplyr::summarise(total_accident=n())

ggplot(accident_weekdays, 
       aes(hours, 
           total_accident, 
           color='red'))+
geom_point()+
geom_line()+
ggtitle('Number of Accidents on Weekdays')

```

#Visualizing weekday trend for each month

```{r}
month_weekday <- accidents_df%>%
                  group_by(month,crash_day_of_week)%>%
                  dplyr::summarise(count=n())
 

month_weekday <- month_weekday[-c(14,72),]


month_weekday$crash_day_of_week <- factor(month_weekday$crash_day_of_week,
  levels = c("SU", "MO", 
  "TU", "WE", "TH", "FR", "SA"))


ggplot(month_weekday, aes(month, 
                          reorder(crash_day_of_week,count),                                        fill=count)) + 
geom_tile() +
scale_fill_gradient(low = "#86ebc9",
                    high = "#09855c",
                    guide = "colorbar") +  
guides(fill=guide_legend(title="Total Accidents")) +
theme_bw() + 
theme_minimal() + 
labs(title = "Histogram of NJ Accidents by Day of Week and Month",
         x = "Accidents by Month", y = "Day of Week") +
theme(panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())

```


##Visualizing number of cases for each day for every hour

```{r}
weekday_hours <- accidents_df%>%
                  group_by(crash_day_of_week, hours)%>%
                  dplyr::summarise(count=n())

weekday_hours <- weekday_hours[-c(1),]

weekday_hours$crash_day_of_week <- factor(weekday_hours$crash_day_of_week,
  levels = c("SU", "MO", 
  "TU", "WE", "TH", "FR", "SA"))
weekday_hours <- weekday_hours[-c(144), ]

ggplot(weekday_hours, aes(hours,
                          crash_day_of_week,
                          fill=count)) + 
geom_tile() +
scale_fill_gradient(low = "#86ebc9",
                    high = "#09855c",
                    guide = "colorbar") +  
guides(fill=guide_legend(title="Total Incidents")) +
theme_bw() + 
theme_minimal() + 
labs(title = "Histogram of NJ Accidents by Day of Week and Hour",
         x = "Accidents Per Hour", y = "Day of Week") +
theme(panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())

```



##Fatality by Year

```{r}
accidents_df <- accidents_df %>% mutate(killed = case_when(total_killed==0 ~ "No", TRUE ~ "Yes"))

fatality <- accidents_df%>%
  group_by(year, killed)%>%
  dplyr::summarise(count=n())

ggplot(data=fatality[fatality$killed %in% c('Yes'),], 
        aes(x=year, y=count, ))+
  geom_bar(stat='identity', fill='lightblue', width=0.3)+
  geom_line(group=1)+
  xlab("Year")+
  ylab("Number of Deaths")+
  coord_cartesian(ylim = c(0, 900))+
  ggtitle("Total Number of People Died each year")

```



```{r}
injured <- accidents_df %>% 
              mutate(injured = case_when(total_injured==0 ~ "No",
                                            TRUE ~ "Yes"))%>%
             group_by(year, injured)%>%
             dplyr::summarise(count=n())


ggplot(injured[injured$injured %in% c('Yes'),], 
        aes(x=year, y=count))+
  geom_bar(stat='identity', fill='lightblue', width=0.3)+
  geom_line(group=1)+
  xlab("Year")+
  ylab("Number of People Injured")+
  ggtitle("Total Number of People Injured each year")

```


##Crashes by Age

```{r}
accidents_df$age <- floor(as.numeric(difftime(Sys.Date(),accidents_df$driver_dob, units = "weeks"))/52.25)



accidents_df%>%
  group_by(year, age)%>%
  filter(age_group != '')%>%
  dplyr::summarise(count=n())%>%
  ggplot(aes(age,count, color=year))+
  geom_line(size=1)+
  coord_cartesian(xlim = c(16, 100))+
  ggtitle('Number of Accidents by Age for each Year' )
```


##Number of crashes by age, sex, and year

```{r}
age_sex_count <- accidents_df%>%
                 group_by(year, age, driver_sex)%>%
                  dplyr::summarise(total_accidents=n())%>%
                  filter(driver_sex !='')

  


age_sex_count %>% mutate(
    total_accidents = ifelse(driver_sex=="M", 
                             total_accidents*(-1),
                        total_accidents*1))%>%
    ggplot(aes(age,total_accidents, fill=driver_sex))+ 
    geom_bar(stat = "identity")+
    coord_flip()+
    facet_wrap(~year,ncol=2)+
    labs(title = "Number of Accidents by Age and Sex", x = "Age",
        y = "Number of Accidents")
    
```



#Number of crashes based on environmental condition

```{r}
accidents_df%>%
  group_by(month, environmental_condition)%>%
  dplyr::summarise(count=n())%>%
  filter(environmental_condition != 99)

accidents_df%>%
  group_by(month, environmental_condition)%>%
  dplyr::summarise(count=n())%>%
  filter(environmental_condition != 99)%>%
  ggplot(aes(x=environmental_condition, y=count, color=month))+
  geom_line()+
  coord_cartesian(xlim = c(0, 10))+
    ggtitle('Number of Accidents based on environmental condition')

```


##Number of crashes in top 10 route

```{r}
accident_by_route <- accidents_df%>%
                      group_by(route)%>%
                      dplyr::summarise(Number_of_Accidents=n())%>%
                      arrange(desc(Number_of_Accidents))%>%
                      slice(1:10)

ggplot(accident_by_route,
       aes(route,
          Number_of_Accidents))+
geom_bar(stat='identity')+
ggtitle('Number of Accidents per Route')
```



#Number of crashes for each intersection

```{r}
accidents_df%>%
  group_by(year, intersection)%>%
  dplyr::summarise(total_accidents=n())%>%
  ggplot(aes(year,
             total_accidents, 
             fill = intersection))+
  geom_bar(stat='identity', beside=TRUE)+
  ggtitle('Accidents at each Intersection')
```



##Number of crashes based on cell phone usage

```{r}
accidents_df%>%
  group_by(year, cell_phone_in_use_flag)%>%
  dplyr::summarise(accidents=n())%>%
  filter(cell_phone_in_use_flag=='Y')%>%
  ggplot(aes(year, 
             accidents,                        
             fill=cell_phone_in_use_flag))+
  geom_bar(stat='identity', fill='lightblue', width=0.3)+
  geom_line(group=1)+
  ggtitle('Number of Accidents ~ Cell Phone Usage')
```


###Number of vehicles involved in crashes

```{r}
accidents_df%>%
  group_by(total_vehicles_involved)%>%
  dplyr::summarise(Number_of_accidents=n())%>%
  filter(total_vehicles_involved <= 6)%>%
  ggplot(aes(total_vehicles_involved, 
             Number_of_accidents,                             
             fill = total_vehicles_involved))+
  geom_bar(stat='identity')+
  ggtitle('Number of Casualties per Accidents')
```


##Number of crashes based on crash code

```{r}
crash_type <- accidents_df%>%
              group_by(crash_type_code)%>%
              dplyr::summarise(total_accident=n())

crash_type <- crash_type[-c(1, 17, 18),]

ggplot(crash_type, 
       aes(reorder(crash_type_code,total_accident),                             total_accident, 
       fill=crash_type_code))+
geom_bar(stat='identity')+
xlab('Crash Code')+
ggtitle('Total Number of Accidents based on Crash Code')


```



##Number of crashes for each day with alcohol consumption

```{r}
alcohol_percentage <- accidents_df%>%
                       group_by(crash_day_of_week, 
                                alcohol_involved)%>%
                       dplyr::summarise(accidents=n())

alcohol_percentage <- alcohol_percentage[-c(1,14),]


alcohol_percentage$percentage <- alcohol_percentage$accidents/sum(alcohol_percentage$accidents)*100

alcohol_percentage$crash_day_of_week <-                                                             ordered(alcohol_percentage$crash_day_of_week,                                    levels=c("MO", "TU", "WE",                                                        "TH", "FR", "SA",                                                        "SU"))

ggplot(alcohol_percentage[alcohol_percentage$alcohol_involved %in%     c('Y'), ],
       aes(percentage,
           reorder(crash_day_of_week, crash_day_of_week), 
           fill=percentage))+
geom_bar(stat='identity')+
ylab('Weekday')+
ggtitle('Percentage of crash by Day of Week with Alcohol')

```
```{r}
acc_20$latitude <- ave(acc_20$latitude,acc_20$municipality_name,FUN=function(x) 
  ifelse(is.na(x), mean(x,na.rm=TRUE), x))

acc_20$longitude <- ave(acc_20$longitude,acc_20$municipality_name,FUN=function(x) 
  ifelse(is.na(x), mean(x,na.rm=TRUE), x))

```


##Density of crashes by county

```{r}
acc_20$longitude <- abs(acc_20$longitude)
acc_20$longitude <- acc_20$longitude * -1

acc_20state <- 'NJ'

county_map_data<-map_data("county")
state_map <- map_data("state")
nj_df <- subset(state_map, region == "new jersey")

nj_base <- ggplot(data = nj_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  theme_void()+
  geom_polygon(color = "white", fill = NA)

nj_base + geom_point(data = acc_20, aes(x = longitude, y = latitude), inherit.aes = FALSE, color = 'red', size = 0.01)
  
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
